Topological Equivariant Artist Model

### 2021-12-27 slides/main.tex
clean up intro, new slide summing math

\documentclass[xcolor={dvipsnames}, handout]{beamer}
%\documentclass[xcolor={dvipsnames}]{beamer}
%\usepackage{amsmath,amsfonts,amssymb,pxfonts,eulervm,xspace}
\usepackage{amsmath, amsfonts, amssymb, mathtools, eulervm, xspace}
\usepackage{bm}
\usepackage{mathrsfs} % math script fonts
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\newtheorem{prop}{Proposition} %math?

@@

\begin{frame}{Visualizations are structure preserving maps}
    \begin{description}
        \item [continuity] how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
        \item [equivariance] data transformations have an equivalent effect on the graphical representation, e.g. rotating a matrix has a matching rotation of the image, translating the points on a line has a matching visual shift in the line plot
      \end{description}
\end{frame}

@@

\begin{frame}{Expressed in Indexing}
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
        \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
    \end{figure}
\end{frame}

\begin{frame}{Why is indexing not enough?}
    \includegraphics[width=1\linewidth]{../paper/figures/whycontinuity.png}
\end{frame}

\begin{frame}{Fiber bundle as abstract data structure\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
\begin{columns}
    \column{0.49\columnwidth}
        \begin{equation}
            \label{eq:related-work:continuity:fiber-bundle}
            \begin{tikzcd}[ampersand replacement=\&, row sep=large]
                \dfiberc \arrow[r, hook] & \dtotalc \arrow[d, "\pi"'] \\
                            & \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
            \end{tikzcd}
        \end{equation}
        %\begin{equation}
        %    \label{eq:related-work:continuity:sections}
        %    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \big\}
        %\end{equation}
 \column{0.49\columnwidth}
\end{columns}  
\end{frame}

\begin{comment}
\begin{frame}{Equivariance}
    \begin{description}
        \item[Retinal Variables \& Marks] visual encodings match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
        \pause
        \item[Expressiveness] structure preserving mappings\cite{mackinlayAutomatingDesignGraphical1986}
        \pause
        \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
        \pause 
        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
    \end{description}
\end{frame}

@@

\subsection{Contributions}
\begin{frame}{Contributions}
    \begin{description}
        \item[Topological] continuity  
        \item[Equivariant] monoid action 
        \item[Artist] Matplotlib $\textit{artist}: \textit{data} \rightarrow \textit{graphic}$ 
        \item [Model] 
    \end{description}
\end{frame}

@@

    def view(self, axes):
        # walk the edge_vertex table to return the edge function
        table = defaultdict(list)
        for (i, (start, end)) in sorted(zip(self.ids, self.vertices), 
                                            key=lambda v:v[1][0]):
            table['index'].append(i)
            # same as view for line, returns nested list
            for (name, value) in zip(self.FB.F.keys(), self.tau(i)[1]):
                table[name].append(value)
        return table
\end{minted}
\end{frame}
\end{comment}
\end{document}

>>>>

\documentclass[xcolor={dvipsnames}, handout]{beamer}
%\documentclass[xcolor={dvipsnames}]{beamer}
%\usepackage{amsmath,amsfonts,amssymb,pxfonts,eulervm,xspace}
\usepackage{amsmath, amsfonts, amssymb, mathtools, eulervm, xspace}
\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
\usepackage{bm}
\usepackage{mathrsfs} % math script fonts
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\newtheorem{prop}{Proposition} %math?

@@

\begin{frame}{Visualizations Preserve Structure}
    \begin{description}
        \item [continuity] topological properties \cite{wilkinsonGrammarGraphics2005}, i.e. how elements in a dataset are organized, e.g. discrete rows in a table, networked nodes, pixels in an image, points on a line
        \item [equivariance] data and visual encodings are matched such that transformations have an equivalent effect on data and graphical representations, e.g. rotating a matrix and image, shifting  points on a line and a line graphic
      \end{description}
\end{frame}

@@

\begin{frame}{Expressed in container}    
    \begin{figure}
        \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
        \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
    \end{figure}
\end{frame}

\begin{frame}{Why is container type not enough?}
    \includegraphics[width=1\linewidth]{../paper/figures/whycontinuity.png}
\end{frame}

\begin{frame}{Equivariance}
    \begin{description}
        \item[Retinal Variables \& Marks] visual encodings match properties of the data \cite{bertinSemiologyGraphicsDiagrams2011a}
        \pause
        \item[Expressiveness] structure preserving mappings\cite{mackinlayAutomatingDesignGraphical1986}
        \pause
        \item[Graphical Integrity] graphs show \textbf{only} the data\cite{tufteVisualDisplayQuantitative2001}
        \pause 
        \item[Naturalness] easier to understand when properties match\cite{norman_things_smart}
    \end{description}
\end{frame}

@@

\subsection{Algebraic Topology & Category Theory}

\begin{frame}{Mathematical Data Abstraction}
    \begin{description}
        \item[Fiber Bundles] "unified, dimension-independent framework" that expresses data as the mapping between continuity and fields \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}
        \item[Simplicial Databases] systemic way to apply a schema to a fiber, i.e. provides a way to name fields and bind them to types system (e.g. \texttt{int}, \texttt{float}) 
         \cite{spivakSIMPLICIALDATABASES, spivakDatabasesAreCategories2010} 
        \item[Sheaves on Bundles] "algebraic data structure" for representing data over topological spaces \cite{ghristElementaryAppliedTopology2014}
    \end{description}
\end{frame}

\begin{frame}{Fiber Bundles}
    \begin{columns}
        \column{0.49\columnwidth}
            \begin{equation*}
                %\label{eq:related-work:continuity:fiber-bundle}
                \begin{tikzcd}[ampersand replacement=\&]
                  \dfiberc \arrow[r, hook] \& \dtotalc \arrow[d, "\pi"'] \\
                   \& \dbasec \arrow[u, "\dsectionc \in \cgamma{\dbasec}{\dtotalc}"', bend right, pos=.5]
                \end{tikzcd}
            \end{equation*}
     \column{0.49\columnwidth}
        \begin{description}
            \item[\textcolor{base}{Base Space}] $( \dbasec, \mathcal{T})$
            \item[\textcolor{fiber}{Fiber Space}] $\dfiberc_{\dbasepointc} = \pi^{-1}(\dbasepointc)$
            \item[\textcolor{total}{Total Space}]  $\dtotalc = \dbasec \times \dfiberc$
        \end{description}
    \end{columns}
    \pause
    \begin{block}{Sections of the bundle}
        \begin{equation*}
            \label{eq:related-work:continuity:sections}
            \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \subset \dbasec \big\}
        \end{equation*}
    \end{block}
\end{frame}
\subsection{Contributions}
\begin{frame}{Contributions}
    \begin{description}
        \item[Topological] continuity  
        \item[Equivariant] monoid action 
        \item[Artist] Matplotlib $\textit{artist}: \textit{data} \rightarrow \textit{graphic}$ 
        \item [Model] 
    \end{description}
\end{frame}

@@

    def view(self, axes):
        # walk the edge_vertex table to return the edge function
        table = defaultdict(list)
        for (i, (start, end)) in sorted(zip(self.ids, self.vertices), 
                                            key=lambda v:v[1][0]):
            table['index'].append(i)
            # same as view for line, returns nested list
            for (name, value) in zip(self.FB.F.keys(), self.tau(i)[1]):
                table[name].append(value)
        return table
\end{minted}
\end{frame}
%\end{comment}
\end{document}
